### 2016-02-25 paper.tex
Scaffolding for SPARK16.

\def\year{2016}\relax
%File: formatting-instruction.tex
\documentclass[letterpaper]{article}
\usepackage{aaai16}
\usepackage{times}
\usepackage{helvet}
\usepackage{courier}

\usepackage[pdftex]{hyperref}
\usepackage{amsmath}
\usepackage{color}
\usepackage{textcomp}
\usepackage{booktabs}
\usepackage{ccicons}
\usepackage{todonotes}
\usepackage{amsfonts}

\frenchspacing
\setlength{\pdfpagewidth}{8.5in}
\setlength{\pdfpageheight}{11in}
\pdfinfo{
/Title (Insert Your Title Here)
/Author (Put All Your Authors Here, Separated by Commas)}
\setcounter{secnumdepth}{0}

\newcommand{\aux}{\mathit{aux}}
\newcommand{\dist}{\mathit{dist}}
\newcommand{\cA}{\mathcal{A}}
\newcommand{\tit}[1]{\textit{#1}}
\newcommand{\IFF}{\textit{iff}}
\newcommand{\IF}{\textit{if}}
\newcommand{\PCF}{\mathit{PCF}}
\newcommand{\walk}{\mathit{walk}}
\newcommand{\bike}{\mathit{bike}}
\newcommand{\car}{\mathit{car}}
\newcommand{\public}{\mathit{public}}
\newcommand{\taxi}{\mathit{taxi}}
\newcommand{\gas}{\mathit{gas}}
\newcommand{\wait}{\mathit{wait}}
\newcommand{\SUM}{\mathit{sum}}
\newcommand{\MAX}{\mathit{max}}
\newcommand{\MIN}{\mathit{min}}
\newcommand{\AVG}{\mathit{avg}}

\newcommand{\secref}[1]{Section~\ref{sec:#1}}
\newcommand{\figref}[1]{Figure~\ref{fig:#1}}
\newcommand{\tblref}[1]{Table~\ref{tbl:#1}}
\newcommand{\algref}[1]{Algorithm~\ref{alg:#1}}
\newcommand{\thmref}[1]{Theorem~\ref{thm:#1}}
\newcommand{\propref}[1]{Proposition~\ref{prop:#1}}
\newcommand{\corref}[1]{Corollary~\ref{cor:#1}}

 \begin{document}
% The file aaai.sty is the style file for AAAI Press 
% proceedings, working notes, and technical reports.
%
\title{On Extensibility and Personalizability of Multi-Modal Trip Planning}
\author{AAAI Press\\
Association for the Advancement of Artificial Intelligence\\
2275 East Bayshore Road, Suite 160\\
Palo Alto, California 94303\\
}
\maketitle
\begin{abstract}
	In the setting of multi-modal trip planning, it is essential for
	the planner to allow users to upload auxiliary data of interest,
	let them express personal constraints and preferences,
	and compute optimal trips accordingly.
	In this work, we designed and implemented a multi-modal trip
	planning framework that is both extensible and personalizable.
	%We conducted a user study on the feasibility of the features of
	%our planner, and showed that the features are desirable and have
	%great potential value in mobility marketplace. 
	To this end, we developed a hipergraph-based trip planner that 
	accommodates constraints specified as a theory in the 
	\tit{linear temporal logic} and preferences as a 
	user-specific \tit{preferential cost function} 
	to compute optimal routes using the A* algorithm.
\end{abstract}

\section{Introduction}
%% \begin{itemize}
%% 	\setlength\itemsep{1pt}
%% 	\item What's the problem: to design and implement a trip planning
%% 				system that is multi-modal, extensible and personalizable.
%% 	\item Why is it important: a story-telling example here should suffice.
%% 	\item Why is it hard: not straightforward how to represent and reason
%% 				about user constraints and preferences, etc.
%% 	\item How did we solve it: allowing the user to upload data and specify
%% 				constraints and preferences as LTL formulas and preferential
%% 				cost functions, respectively.  Then, these constraints and
%% 				preferences are integrated into the A* search.
%% 	\item How did we evaluate the features and what are the results:
%% 				ceteris paribus experimentation setup on AMT and results
%% \end{itemize}

\begin{itemize}
	\item What's the problem: personalized trip planning: find trips that
	  are optimal in the eyes of a user, who doesn't care about mode or
	  provider boundaries, and whose preferences may be more faceted than
	  ``fastest'' or ``cheapest''.
	\item Why is it hard: the computational complexity of searching
	  combinations of multiple modes that can be combined freely is high
	  (PSPACE-complete\cite{}) which is why existing trip planner exploit problem
	  structure (such as only one mode per trip) and predefined
	  preferences (optimization metrics)
	\item How did we solve it: Using a high-performance graph search
	  engine\cite{}, we implemented a multi-modal trip planner that uses pure
	  graph-search. This allows us to flexibly combine various modes
	  (e.g., car, train, cab, walk) and declarative change soft and hard
	  constraints. The planner can take constraints expressed in linear
	  temporal logic to constrain the search space (e.g., never bike after
	  transit, i.e., I don't want to take my bike on the bus or train).
	  The planner uses a weighted sum over several metrics (e.g., time
	  overall, time spend on transit, cost) which can be re-weighted based
	  on different user preferences. Finally the planner allows the user
	  to upload {\em new} mapping data over which to express hard and/or
	  soft constraints. For instance, a user might upload a map of crime
	  in the city, and ask the trip planner to never walk at night through
	  an area where crime is frequent.
%	\item How did we evaluate the features and what are the results:
%	  ceteris paribus experimentation setup on Amazon Mechanical 
%		Turk (MTurk\footnote{\url{http://mturk.amazon.com}}) and results, 
%		asking users to pick between two different trips, one being limited by the
%	  capabilities of existing planners, and one exhibiting features only
%	  available in our trip planner.
\end{itemize}

\section{Related Work}
\begin{itemize}
	\setlength\itemsep{1pt}
	\item Summarize related work on multi-modal trip planning.
	\item Summarize related work on constraints and preferences in planning.
		\begin{itemize}
			\setlength\itemsep{0pt}
			\item LTL in planning.
			\item Numeric/ordinal/qualitative preferences in planning.
		\end{itemize}
	\item Summarize related work on VOT (Value-of-Time).
\end{itemize}
Constraints and preferences in planning have been studied in the 
literature.
Son and Pontelli presented a declarative preference language, 
$\mathcal{PP}$, to represent ordinal preferences (e.g., basic desires,
atomic preferences and general preferences) between trajectories of 
a planning problem~\cite{son2004planning}.
Bienvenu et al. proposed a first-order preference language, $\mathcal{LPP}$, 
extending $\mathcal{PP}$ by allowing the specification of quantified qualitative
preferences~\cite{bienvenu2011specifying}.
Bast et al. surveyed algorithms, including Dijkstra and A*, in the setting of
multi-modal transportation networks ~\cite{bast2015route}.
Modesti and Sciomachen introduced an ad hoc utility function for weighing
the arcs both with their time and monetary cost~\cite{modesti1998utility}.
Economic models capturing the trade-off between travel time and travel
cost have been presented~\cite{antoniou2007methodology}.

\section{Extensibility}
\begin{itemize}
	\setlength\itemsep{1pt}
	\item Explain in detail what extensibility means.
	\item Describe the input.
	\item Describe how we incorporate the auxiliary data into the planner.
\end{itemize}
Allowing users to upload their own data sets of interests is an important
step towards customization of a trip planner.
We designed a framework where a user can load the data sets (e.g., crime
statistics and pollution data) into the planner,
and the planner will compute an optimal route accordingly.

\subsection{Auxiliary Data}
The user-created data are the auxiliary data that is represented as pairs
of latitude and longitude degrees.
To merge these lat-long pairs into the planner, we performed a neighborhood
search to calculate the total score of auxiliary data for each 
lat-long pair already in our planner.
It might be of strong interest to some user for our planning
system to take care of criminal statistics so that some level of safety
of the resulting routes is guaranteed.
For instance, a user traveling through the downtown area of San Francisco
around midnight may want to upload a data set of crimes (cf. \figref{sf_crime}), 
and express her constraints and preferences in hope of a preferred trip plan.

\begin{figure}[!ht]
  \centering
    \includegraphics[width=0.4\textwidth]{figs/sub_sf_crime_July2015.png}
  \caption{Crime rates in San Francisco\label{fig:sf_crime}}
\end{figure}

Formally, let $A$ be the auxiliary set of lat-long pairs, 
$N$ a node in our planner and $r$ the effective radius.
We first compute the set $P$ of auxiliary points such that
$P \subseteq A$ and $\dist(p,N)<r$ for all $p \in P$.
Then, the total score of auxiliary data for $N$ with respect to $P$ and $r$,
denoted by $S_\aux(N,P,r)$, is computed as follows.

\begin{equation}
	S_\aux(N,P,r) = \sum_{\substack{p \in P}} \Big(1 - \frac{\dist(p,N)}{r}\Big)
\end{equation}

\section{Personalizability}
Personalizability consists of two aspects: constraints and preferences.
From the viewpoint of the planner,
constraints, also referred to as hard constraints, are statements that the planner
has to satisfy during the planning process; whereas preferences, also called
soft constraints, are specifications that the planner will need to optimize.
We formulated constraints using linear temporal logic (LTL) and preferences as
a preferential cost function (PCF), and implemented our planner leveraging the
widely-used graph search algorithm the A*.

\subsection{Constraints}
\begin{itemize}
	\setlength\itemsep{1pt}
	\item Describe syntax and semantics of LTL in the setting of trip planning.
	\begin{itemize}
		\setlength\itemsep{0pt}
		\item Specifically, we need to point out the ``after" temporal connective,
					and we need to describe the semantics: what is a plan, why we can
					eliminate actions and put them as state attributes instead, etc.
	\end{itemize}
\end{itemize}
As constraints in the setting of trip planning are often declarative and
temporal, our choice of LTL is straightforward.
We now give a brief review of linear temporal logic (LTL).
Let $f$ be a propositional formula over a finite set $L$ of Boolean variables.  
LTL formulas are defined recursively as follows.
\begin{equation}
	\varphi = f | \varphi_1 \land \varphi_2 | \varphi_1 \lor \varphi_2 | \neg \varphi | 
		\bigcirc \varphi |	\Box \varphi | \Diamond \varphi | \varphi_1 \cA \varphi_2
\end{equation}
Note that we have $\varphi_1 \cA \varphi_2$, and it means that
``$\varphi_2$ holds right after $\varphi_1$ holds."

An intuitive constraint in trip planning could be ``I will never drive after biking or
taking the public transit."
In LTL, such constraint can be translated into an LTL formula
\begin{align*}
	\psi = ((M=b) \lor (M=p)) \,\cA\, (\Box (\neg (M=c))).
\end{align*}

As the actions in trip planning is limited to taking different transportation modes,
in our definition of the semantics of LTL
these actions are subsumed into the interpretations of $L$, or \tit{states}.
The semantics of LTL is defined with regard to trajectories of states. 
Let $\sigma$ be a trajectory of states $S_0,a_1,S_1,\ldots,a_n,S_n$, and
$\sigma[i]$ a suffix $S_i, a_{i+1}, S_{i+1}, \ldots,a_n,S_n$.  We have
\begin{align*}
	\sigma \models f \;\; &\IFF \;\; S_0 \models f,\\
	\sigma \models \varphi_1 \land \varphi_2 \;\; &\IFF \;\; \sigma \models \varphi_1 \; and \; \sigma \models \varphi_2,\\
	\sigma \models \varphi_1 \lor \varphi_2 \;\; &\IFF \;\; \sigma \models \varphi_1 \; or \; \sigma \models \varphi_2,\\
	\sigma \models \neg \varphi \;\; &\IFF \;\; \sigma \not \models \varphi,\\
	\sigma \models \bigcirc \varphi \;\; &\IFF \;\; \sigma[1] \models \varphi,\\
	\sigma \models \Box \varphi \;\; &\IFF \;\; \forall 0 \leq i \leq n (\sigma[i] \models \varphi),\\
	\sigma \models \Diamond \varphi \;\; &\IFF \;\; \exists 0 \leq i \leq n (\sigma[i] \models \varphi),\\
	\sigma \models \varphi_1 \cA \varphi_2 \;\; &\IFF \;\; \forall 0 \leq i < n (\IF \; \sigma[i] \models \varphi_1, \sigma[i+1] \models \varphi_2).
\end{align*}

For example, we are given an LTL constraint $\psi$ and three trajectories ($\sigma_1$, $\sigma_2$ and $\sigma_3$)
as shown in \figref{trjs}.
Clearly, we have $\sigma_1 \models \psi$ because, after public transit in $S_2$ and $S_3$, 
traveling by car has never been taken place.
Moreover, we have $\sigma_2 \not \models \psi$ because we have $M=c$ hold in $S_6$ and $S_7$
after having $M=p$ hold in $S_2$ and $S_5$.
Finally, we know $\sigma_3 \models \psi$, as the mode is always neither biking nor public transit.

\begin{figure}[!ht]
  \centering
    \includegraphics[width=0.5\textwidth]{figs/trajectories.pdf}
  \caption{State transition diagram\label{fig:trjs}}
\end{figure}

\subsection{Preferences}
\begin{itemize}
	\setlength\itemsep{1pt}
	\item Describe our preferential cost function (PCF).
	\begin{itemize}
		\setlength\itemsep{0pt}
		\item Describe why and how we break down time and money.
		\item Describe why and how we design coefficients to these time and money pieces.
		\item Describe why and how we design the two ratios: dollars per hour and dollars per aux.
	\end{itemize}
\end{itemize}

A state is described as a set of \tit{state variables}.
The state variables of a state $S$ include the transportation mode $M$ that led to $S$,
time spent $T$ so far per mode (e.g., $T_\bike$ for biking and $T_\public$ for
public transit), fare $D$ spent so far per mode (e.g., $D_\gas$ for driving and
$D_\taxi$ for taking a cab), and variables related to the auxiliary data once uploaded.
These extra data related variables are metrics such as the sum ($A_\SUM$),
the maximum ($A_\MAX)$, the minimum ($A_\MIN$), and the average ($A_\AVG$) data along the path.

We focused on weighted functions over state variables and
designed the cost function, called preferential cost function (PCF), that guides the
graph-based search engine in our trip planner as follows.
\begin{equation}
	\begin{aligned}
		\PCF(S) =& \beta_1 * (\alpha_1 \cdot T_\walk + \alpha_2 \cdot T_\wait + \ldots) \\
								&+ (D_\gas + D_\public + \ldots) \\
								&+ \beta_2 * (A_\SUM + \ldots),
	\end{aligned}
\end{equation}
where $\alpha_i$ are real numbers representing the relations among different time pieces,
and $\beta_1$ ($\beta_2$) is the ratio that essentially describes how much in dollars a user would pay to
save an hour (an auxiliary data, respectively).

\subsubsection{Preference Elicitation}
To gather these coefficients ($\alpha_i$ and $\beta_i$) in our $\PCF$, we designed interface to
elicit these numbers from the user.

\subsection{Reasoning with Constraints and Preferences}
\begin{itemize}
	\setlength\itemsep{1pt}
	\item Describe why and how we integrate constraints and preferences into A*.
\end{itemize}

We leveraged the widely-used A* search algorithm on top of our high-performance graph search
engine.  The A* algorithm incorporates the following cost function.
\begin{equation}
	f(S) = g(S) + h(S),
\end{equation}
where $g(S)$ is the cost of an optimal trip from the initial state to $S$, and
$h(S)$ is an admissible estimate of the cost of an optimal trip from $S$ to goal.

To prune the search space, we check satisfiability of the temporal constraints in LTL
at expansion of the search tree.
To guide the search engine, we set $g(S)=\PCF(S)$ and $h(S)$ the minimum estimate among
all available modes in $S$.

%\section{User Study}
%\begin{itemize}
%	\setlength\itemsep{1pt}
%	\item Describe the set up of our experiments on AMT and explain why: because we want to evaluate the
%				feasibility of the three features.
%	\item Describe the results.
%\end{itemize}
\section{Implementation}

\begin{itemize}
	\setlength\itemsep{1pt}
	\item Describe the system structure.
	\item Describe the input to the system.
	\item Describe the output from the system.
\end{itemize}

\section{Conclusion and Future Work}

\bibliographystyle{aaai}
\bibliography{refs}

\end{document}