### 2016-03-03 summary.tex

In multi-modal trip planning, it is vital for the planner to enable
users to integrate new metrics of interest into the existing map,
and to express constraints and preferences over which optimal
routes are computed.
In our work, we designed and built a graph-search based, extensible
and personalizable multi-modal trip planner that utilizes linear
temporal logic and preferential cost
function to model user constraints and preferences.

As a note on the future work, we plan to explore techniques in planning
for computing multiple routes that are diverse with bounded difference of costs.
Continuing this direction, we intend to study the problem of learning the
$\PCF$ coefficients using the observations of the decisions the user made among
the computed paths. (It is reasonable to assume that routes picked by the user
are preferred to others, and these observations could contribute to the learning
problem.)
Also interesting is to introduce traffic information into the planner
to support real-time multi-agent concurrent trip planning.